---
title: "Replication of 'Sexism and the Far-Right Vote: The Individual Dynamics of
  Gender Backlash' (Anduiza & Rico, 2022)"
author: "Reeka Estacio and Belynda Herrera"
date: "2025-03-05"
output: pdf_document
---

```{r setup, include=FALSE}
# hide code chunks in rendered document
knitr::opts_chunk$set(echo = FALSE) 
library(tidyverse)
```

# Introduction
The study "Sexism and the Far-Right Vote: The Individual Dynamics of Gender Backlash" by Eva Anduiza and Guillem Rico, published in the American Journal of Political Science (AJPS), examines the relationship between sexist attitudes and support for far-right political parties. The authors argue that gender backlash—a reactionary response to perceived advances in gender equality—plays a crucial role in shaping voting behavior.

Using survey data collected between 2017 and 2020, the study investigates whether changes in sexism over time are associated with increased support for Vox, a far-right party in Spain. The authors hypothesize that individuals who express greater sexist attitudes over time are more likely to support the far-right, reflecting a broader trend of gendered political polarization.


## Original Study

```{r}
vox <- read.csv("data/cleaned.csv")
head(vox)
```


What is the unit of analysis? (what are the cases in the dataset)
The unit of analysis in this dataset is individual survey respondents. Each row in the dataset represents an individual who participated in the survey, providing responses on their political preferences, sexist attitudes, and demographic characteristics over multiple time periods.

Any concerns about observational independence?
Observational independence concerns exist because the longitudinal dataset (2017-2020) introduces within-subject correlation as the same individuals were tracked over time. Respondents may modify their self-reported sexism and political preferences based on social norms, creating measurement bias. Additionally, major events like #MeToo and Women's Day protests likely influenced responses across the sample simultaneously, creating time-based dependencies.

How many observations are there?

How was the data collected?
The dataset was collected through survey responses in Spain, covering multiple waves from 2017 to 2020. The data was collected before and after key political events (e.g., elections, feminist protests) to capture potential shifts in attitudes and voting behavior.

### Original Model 

What is the dependent variable? Describe it, what kind of data is it? (Binary)
The dependent variable measures whether a respondent votes for a far-right party or not. It captures voting behavior as an outcome influenced by sexist attitudes and other variables.
**PLOT THE DISTRIBUTION OF THE DV**

What are the predictor variables?
The study examines several predictors to explain support for far-right parties, with a focus on sexist attitudes as the central explanatory variable.

What kind of model are you replicating? (logit)

What do the original authors hope to achieve with this model?

**INCLUDE SCREENSHOT OF ORIGINAL REGRESSION TABLE**

# Replication of Original Model

As in the original model, we fit a logistic regression on...

The regression table is presented below.

```{r}
# Replicate model

# modelsummary(replicated_model, output="gt")
```

# Additional Model

We also considered (adding new predictor/removing predictor/changing link function) because...

```{r}
# Additional model

# modelsummary(new_model, output="gt")
```

# Comparing Models

## In-Sample Prediction

```{r}
# Show both models

# models <- list(original_model, new_model)
# modelsummary(models, output="gt")
```

Compare AIC/BIC, log likelihood, MSE.

## Out-of-Sample Prediction

```{r}
# Use cross-validation
# Calculate error
```

Interpret this. 

Based on the error, (original/new) model performs better on out-of-sample prediction.

Ultimately, we believe the **(original/new) model** is the better model because...

# Exploring the Effect of (Quantity of Interest)

We aim to explore the effect of (quantity of interest) on (dependent variable).

```{r}
# From problem set
# ADAPT TO CURRENT DATA

# Make scenarios
# Hold sanctions75 and pop75 fixed at the median 
scenarios <- data.frame(
  sanctions75 = median(dat$sanctions75),
  pop75 = median(dat$pop75),
  polrights75 = seq(min(dat$polrights75, na.rm = TRUE), max(dat$polrights75, na.rm = TRUE)) # compare polrights75 at different values
)

# Generate predictions
predictions <- predict(nb_mod, newdata = scenarios, type = "response", se.fit = TRUE)

# Add predictions and confidence intervals to the dataframe
scenarios$predicted_deaths75 <- predictions$fit
scenarios$lower_ci <- predictions$fit - 1.96 * predictions$se.fit  
scenarios$upper_ci <- predictions$fit + 1.96 * predictions$se.fit 

# Plot predictions
ggplot(scenarios, aes(x = polrights75, y = predicted_deaths75)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), alpha = 0.2, fill = "blue") +
  labs(title = "Effect of polrights75 on predicted deaths in 1975",
       x = "`polrights75`",
       y = "Predicted deaths in 1975") +
  theme_minimal()
```

Interpret the results.

Based on this plot, we see that (quantity of interest) has a ____ relationship on (dependent variable).

# Summary

Summarize conclusions.

How confident are you about the original author's conclusions?

What else would you like to see done with this?

# Appendix

We used ChatGPT to...

[Link to conversation]()
