---
title: "Replication of 'Sexism and the Far-Right Vote: The Individual Dynamics of
  Gender Backlash' (Anduiza & Rico, 2022)"
author: "Reeka Estacio and Belynda Herrera"
date: "2025-03-05"
output: pdf_document
---

```{r setup, include=FALSE}
# hide code chunks in rendered document
knitr::opts_chunk$set(echo = FALSE) 
library(tidyverse)
```

# Introduction

## Original Study

What is the unit of analysis? (what are the cases in the dataset)

Any concerns about observational independence?

How many observations are there?

How was the data collected?

### Original Model 

What is the dependent variable? Describe it, what kind of data is it? (Binary)
**PLOT THE DISTRIBUTION OF THE DV**

What are the predictor variables?

What kind of model are you replicating? (logit)

What do the original authors hope to achieve with this model?

**INCLUDE SCREENSHOT OF ORIGINAL REGRESSION TABLE**

# Replication of Original Model

As in the original model, we fit a logistic regression on...

The regression table is presented below.

```{r}
# Replicate model

# modelsummary(replicated_model, output="gt")
```

# Additional Model

We also considered (adding new predictor/removing predictor/changing link function) because...

```{r}
# Additional model

# modelsummary(new_model, output="gt")
```

# Comparing Models

## In-Sample Prediction

```{r}
# Show both models

# models <- list(original_model, new_model)
# modelsummary(models, output="gt")
```

Compare AIC/BIC, log likelihood, MSE.

## Out-of-Sample Prediction

```{r}
# Use cross-validation
# Calculate error
```

Interpret this. 

Based on the error, (original/new) model performs better on out-of-sample prediction.

Ultimately, we believe the **(original/new) model** is the better model because...

# Exploring the Effect of (Quantity of Interest)

We aim to explore the effect of (quantity of interest) on (dependent variable).

```{r}
# From problem set
# ADAPT TO CURRENT DATA

# Make scenarios
# Hold sanctions75 and pop75 fixed at the median 
scenarios <- data.frame(
  sanctions75 = median(dat$sanctions75),
  pop75 = median(dat$pop75),
  polrights75 = seq(min(dat$polrights75, na.rm = TRUE), max(dat$polrights75, na.rm = TRUE)) # compare polrights75 at different values
)

# Generate predictions
predictions <- predict(nb_mod, newdata = scenarios, type = "response", se.fit = TRUE)

# Add predictions and confidence intervals to the dataframe
scenarios$predicted_deaths75 <- predictions$fit
scenarios$lower_ci <- predictions$fit - 1.96 * predictions$se.fit  
scenarios$upper_ci <- predictions$fit + 1.96 * predictions$se.fit 

# Plot predictions
ggplot(scenarios, aes(x = polrights75, y = predicted_deaths75)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), alpha = 0.2, fill = "blue") +
  labs(title = "Effect of polrights75 on predicted deaths in 1975",
       x = "`polrights75`",
       y = "Predicted deaths in 1975") +
  theme_minimal()
```

Interpret the results.

Based on this plot, we see that (quantity of interest) has a ____ relationship on (dependent variable).

# Summary

Summarize conclusions.

How confident are you about the original author's conclusions?

What else would you like to see done with this?

# Appendix

We used ChatGPT to...

[Link to conversation]()
